---
title: "Homework7"
author: "David Piccirilli"
date: "2023-03-08"
output: 
  html_document: 
    highlight: textmate
    theme: cerulean
---

# Simulating and fitting data distributions
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
library(MASS)
```

### Add Data vector

```{r, eval= F,echo= F}
z <- rnorm(n=3000,mean=0.2) # generate a random data vector with normal distribution
z <- data.frame(1:3000,z) # adding ID numbers 1:3000 called "z"
names(z) <- list("ID","Length") # add column names to data frame "z"
z <- z[z$Length>0,] # subset positive values of Length
str(z)# display z
summary(z$Length)
```

## Sunflower Ovary data set

```{r, echo= F}
z<-read.csv("SunflowerOvary.csv")
str(z)
summary(z)
```

## Plot histogram of data

```{r, echo=T}
# histogram of z and Length
p1 <- ggplot(data=z, aes(x=Length, y=..density..)) +
  geom_histogram(color="black",fill="magenta",linewidth=0.2) 
# x-axis is Length values, y-axis is probability density 
print(p1)
```

## Add density curve

```{r, echo=T}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

## Get maximum likelihood parameters for normal distribution

```{r, echo = T}
normPars <- fitdistr(z$Length,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"]
```

## Plot normal probability density

```{r echo=T, warning=FALSE}
meanML <- normPars$estimate["mean"]# maximum likelihood for normal probability
sdML <- normPars$estimate["sd"] 

xval <- seq(0,max(z$Length),len=length(z$Length))
# the x values are sequenced from the beginning of the data frame to the end of the data frame. length of one column.
 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="green4", n = length(z$Length), args = list(mean = meanML, sd = sdML))
 p1 + stat
 # using the maximum likelihood estimators to draw another curve of the normal distribution
```

## Plot exponential probability density

```{r}
expoPars <- fitdistr(z$Length,"exponential")
rateML <- expoPars$estimate["rate"] #maximum likelihood for exponential probability

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue2", n = length(z$Length), args = list(rate=rateML))
 p1 + stat + stat2
 #continuous exponetial line that does not cross zero (always positive)
```

## Plot uniform probability density

```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkviolet", n = length(z$Length), args = list(min=min(z$Length), max=max(z$Length)))
 p1 + stat + stat2 + stat3
 # does not account for the maximum probability 
 #the line does not fit the data very well and shows teh data in not uniform 
```

## Plot gamma probability density

```{r}
gammaPars <- fitdistr(z$Length,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$Length), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
 # the gamma distribution is a combination of the normal and exponential distribution because the line follows the graph more closely and is the middle values between the lines.
 #takes into account the shape and distribution of the data
```

## Plot beta probability density
 
```{r}
#the beta probability to based on the successes and fails
#the values are only between 0 and 1
pSpecial <- ggplot(data=z, aes(x=Length/(max(Length + 0.1)), y=..density..)) +
  geom_histogram(color="black",fill="magenta",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")
#ploted with a dotted line
betaPars <- fitdistr(x=z$Length/max(z$Length + 0.1),start=list(shape1=1,shape2=2),"beta")
#the success/ the maximum value from the graph
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]
#takes the success and failures (shape1 +1 and shape2 +2)
statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="cyan3", n = length(z$Length), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

## Problem 3

```{r}
# the best fit distribution is the beta probability distibution because the line best fits the data. The graph is not normal distribution and it is not an exponential distribution. 
```

## Problem 4

```{r}
NewBeta<-rbeta(1:77, shape1ML, shape2ML, ncp=0)
d<- data.frame(1:77,NewBeta)
names(d)<-list("ID", "myLength")
str(d)
summary(d$myLength)
```

## New Beta Plot

```{r}
#ploted with a dotted line
betaPars <- fitdistr(x=z$Length/max(z$Length + 0.1),start=list(shape1=1,shape2=2),"beta")


p1 <- ggplot(data=d, aes(x=myLength, y=..density..)) +
  geom_histogram(color="black",fill="magenta",linewidth=0.2) 

print(p1)

betacurve <-stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="cyan3", n = length(d$myLength), args = list(shape1=shape1ML,shape2=shape2ML))

p1 + betacurve
```
